---
title: "Lab 3"
author: "Axel Holmberg (axeho681)"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

# Lab 3

## Assignment 1 - Gaussian kernel

Assignment 1 is about implementing a kernel method for to predict the hourly temperature for a user-defined date and place in Sweden. In the code below the chosen date is $2016-12-24$ and the place is Stockholm defined by the coordinated $59.325$, $18.071$. I implemented three kernels based on the date, time and distance using a gaussian kernel with the gaussian function $e^{-u^2}$.

```{r}
setwd("~/Programming/TDDE01/Lab 3")
RNGversion('3.5.1')
set.seed(1234567890)

library("geosphere")

stations <- read.csv("stations.csv")
temps <- read.csv("temps50k.csv")
st <- merge(stations, temps, by = "station_number")

## These values are up to the user. ud = user defined
ud.lat <- 59.325 # The lat of the point to predict
ud.long <- 18.071 # The long of the point to predict
ud.date <-"2016-12-24" # The date to predict (up to the students) 
ud.h_distance <- 100000
ud.h_date <- 10
ud.h_time <- 4
##End input

ud.times <- c("04:00:00", "06:00:00","08:00:00","10:00:00","12:00:00","14:00:00",
           "16:00:00","18:00:00","20:00:00","22:00:00", "24:00:00")

#Filters posterior date
filterPosteriorDate <- function(data,date) {
  return(data[!(as.Date(data$date)-as.Date(date))>0,])
}

#Filters posterior time
filterPosteriorTime <- function(data, date, time) {
  return (data[!(as.Date(data$date) == as.Date(date) & 
                 as.numeric(difftime(strptime(data$time, format="%H:%M:%S"), 
                 strptime(time, format="%H:%M:%S"))) > 0 ),])
}



#Kernel for distance computing to point of interest
distGaussian <- function(data, target, h) {
  dist <- distHaversine(data.frame(data$longitude,data$latitude), target)
  u <- dist/h
  return (exp(-(u)^2));
}

#Kenrel for date
dateGaussian <- function(data, target, h) {
  date_diff <- as.numeric(as.Date(data$date)-as.Date(target), unit="days")
  u <- date_diff/h
  return(exp(-(u)^2))
}

#Kernel for time
timeGaussian <- function(data,target,h) {
  time_difference <- difftime(strptime(data$time, format="%H:%M:%S"), 
                              strptime(target, format="%H:%M:%S"))
  u <- as.numeric(time_difference)/h
  return(exp(-(u)^2))
}
  
tempEst <- function(data) {
  filtered_data <- filterPosteriorDate(data, ud.date)
  length=length(ud.times) 
  t_sum <- vector(length=length)
  t_multi <- vector(length=length)
  for  (i in 1:length) {
    filtered_data_by_time <- filterPosteriorTime(filtered_data, ud.date, ud.times[i])
    time <- timeGaussian(filtered_data_by_time, ud.times[i], ud.h_time)
    distance <- distGaussian(filtered_data_by_time, data.frame(ud.long, ud.lat), ud.h_distance)
    day <- dateGaussian(filtered_data_by_time, ud.date, ud.h_date)
    kernel_sum <- distance + day + time
    kernel_multi <- distance * day * time
    t_sum[i] <- sum(kernel_sum %*% filtered_data_by_time$air_temperature)/sum(kernel_sum)
    t_multi[i] <- sum(kernel_multi %*% filtered_data_by_time$air_temperature)/sum(kernel_multi)
  }
  
  return(list(t_sum=t_sum, t_multi=t_multi))
}
  
temps <- tempEst(st)

plot(temps$t_sum,xaxt='n', xlab="Time", 
     ylab="Temperature", type="o", main = "Sum of kernels")
axis(1, at=1:length(ud.times), labels=ud.times)

```

The plot above shows the kernel for the sum of the kernels.

``` {r}

plot(temps$t_multi,xaxt='n', xlab="Time", 
     ylab="Temperature", type="o", main = "Product of kernels")
axis(1, at=1:length(ud.times), labels=ud.times)

```

The plot above shows the plot for the product of kernels.

The width chosen for the three kernels are $100 000$ for distance, $10$ for date and $4$ for time. The reasoning for the kernel width is that it gives the most spread out data points for the
## TODO Skriv färdigt detta

## Assignment 2 - Neural network for the sinus function

Assignment 2 is about implementing a neural network for the sinus function based on 50 data points and then choosing the best 

```{r}

library(neuralnet)
library(ggplot2)

set.seed(1234567890)
#50 values between 0 and 10
Var <- runif(50, 0, 10)

# Create dataset
trva <- data.frame(Var, Sin=sin(Var))

# Divide dataset into training and validation set
tr <- trva[1:25,] # Training
va <- trva[26:50,] # Validation

# Random initialization of the weights in the interval [-1, 1]
# 31 weights are used
winit <- runif(31, -1, 1)

# Function predicting MSE
mse <- function(prediction, observation) {
  return (mean((observation - prediction)^2))
}

mse_val <- numeric()
mse_train <- numeric()
threshold <- numeric()
m_sq_err <- function(pred, obs) {
  return (mean((observation - prediction)^2))
}

for(i in 1:10) {
  nn <- neuralnet(Sin ~ Var, data = tr, startweights = winit, hidden = c(10),
                  threshold = i/1000)
  
  pred_train <- compute(nn, covariate=tr)$net.result
  pred_val <- compute(nn, covariate=va)$net.result
  threshold[i] <- i/1000
  mse_val[i] <- mse(pred_val, va$Sin)
  mse_train[i] <- mse(pred_train, tr$Sin)

}

plot(threshold, mse_train, type="o", ylab="MSE", main = "Training dataset")
plot(threshold, mse_val, type="o", ylab="MSE", main = "Validation dataset")

```

The plots above shows the MSE for the validaiton dataset and the training dataset. One can see that the MSE is lowest for $i=4$ in the validation plot, and is thereby chosen for the neural network.

```{r}

# Your code here
nn <- neuralnet(Sin ~ Var, data = trva, startweights = winit, hidden = c(10),
                     threshold = 4/1000)
plot(nn)
```

Above the neural net can be seen with its' 10 hidden nodes and its' acoompanying values.

```{r}
pred_nn <- prediction(nn)$rep1
plot_net <- ggplot() + geom_point(aes(pred_nn[,1], pred_nn[,2])) + 
  geom_point(aes(trva$Var, trva$Sin), colour="red")
plot_net
```

Finally the plot with the predicted data (black points) plotted against the original data(red points). As one can see the points almost perfectly line up, which is the desired outcome.

# Appendices for code

## Appendix 1 - Assignment 1

```{r echo=TRUE, warning=FALSE, eval=FALSE}
setwd("~/Programming/TDDE01/Lab 3")
RNGversion('3.5.1')
set.seed(1234567890)

library("geosphere")

stations <- read.csv("stations.csv")
temps <- read.csv("temps50k.csv")
st <- merge(stations, temps, by = "station_number")

## These values are up to the user. ud = user defined
ud.lat <- 59.325 # The lat of the point to predict
ud.long <- 18.071 # The long of the point to predict
ud.date <-"2016-12-24" # The date to predict (up to the students) 
ud.h_distance <- 100000
ud.h_date <- 10
ud.h_time <- 4
##End input

ud.times <- c("04:00:00", "06:00:00","08:00:00","10:00:00","12:00:00","14:00:00",
           "16:00:00","18:00:00","20:00:00","22:00:00", "24:00:00")

#Filters posterior date
filterPosteriorDate <- function(data,date) {
  return(data[!(as.Date(data$date)-as.Date(date))>0,])
}

#Filters posterior time
filterPosteriorTime <- function(data, date, time) {
  return (data[!(as.Date(data$date) == as.Date(date) & 
                 as.numeric(difftime(strptime(data$time, format="%H:%M:%S"), 
                 strptime(time, format="%H:%M:%S"))) > 0 ),])
}



#Kernel for distance computing to point of interest
distGaussian <- function(data, target, h) {
  dist <- distHaversine(data.frame(data$longitude,data$latitude), target)
  u <- dist/h
  return (exp(-(u)^2));
}

#Kenrel for date
dateGaussian <- function(data, target, h) {
  date_diff <- as.numeric(as.Date(data$date)-as.Date(target), unit="days")
  u <- date_diff/h
  return(exp(-(u)^2))
}

#Kernel for time
timeGaussian <- function(data,target,h) {
  time_difference <- difftime(strptime(data$time, format="%H:%M:%S"), 
                              strptime(target, format="%H:%M:%S"))
  u <- as.numeric(time_difference)/h
  return(exp(-(u)^2))
}
  
tempEst <- function(data) {
  filtered_data <- filterPosteriorDate(data, ud.date)
  length=length(ud.times) 
  t_sum <- vector(length=length)
  t_multi <- vector(length=length)
  for  (i in 1:length) {
    filtered_data_by_time <- filterPosteriorTime(filtered_data, ud.date, ud.times[i])
    time <- timeGaussian(filtered_data_by_time, ud.times[i], ud.h_time)
    distance <- distGaussian(filtered_data_by_time, data.frame(ud.long, ud.lat), ud.h_distance)
    day <- dateGaussian(filtered_data_by_time, ud.date, ud.h_date)
    kernel_sum <- distance + day + time
    kernel_multi <- distance * day * time
    t_sum[i] <- sum(kernel_sum %*% filtered_data_by_time$air_temperature)/sum(kernel_sum)
    t_multi[i] <- sum(kernel_multi %*% filtered_data_by_time$air_temperature)/sum(kernel_multi)
  }
  
  View(distance)
  View(day)
  View(time)
  
  return(list(t_sum=t_sum, t_multi=t_multi))
}
  
temps <- tempEst(st)

plot(temps$t_sum,xaxt='n', xlab="Time", 
     ylab="Temperature", type="o", main = "Sum of kernels")
axis(1, at=1:length(ud.times), labels=ud.times)

plot(temps$t_multi,xaxt='n', xlab="Time", 
     ylab="Temperature", type="o", main = "Multiplication of kernels")
axis(1, at=1:length(ud.times), labels=ud.times)
```

